---
title: 'Class 06: R Functions'
author: "Seth McNeely"
date: "1/23/2020"
output:
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Importing

Let's see more about file import (i.e. reading files into R). The main read function in base R is `read.table()`:
```{r}
t1 <- read.table("test1.txt")
t1
```

We need to add arguments to get this file imported in a proper format:
```{r}
t1 <- read.table("test1.txt", sep=",", header=TRUE)
t1
```

Or I could use `read.csv()` which has these arguments by default!
```{r}
t1 <- read.csv("test1.txt")
t1
```

## Simple functions

Our first example function:
```{r}
add <- function(x, y=1) {
  # Sum the input x and y (y = 1)
  x + y
}
```

This function requires an input "x," but also allows an optional input "y". By default, we gave "y" the value of 1. Now, let's try this function out!
```{r}
# Adding input x to input y
add(7, 3)
```

How does this work with vectors as input?
```{r}
# Adding input vector x to default y
add(c(1,2,4))
```

```{r}
# Adding inout vector x to input y
add(c(1,2,4), 4)
```

```{r}
#Adding input vector x to inout vector y
add(c(1,2,4), c(1,2,4))
```

## Let's move on!
What about this `range()` function?

```{r}
x <- c(1,2,4)
min(x)
max(x)
```
```{r}
range(x)
```

This is our second function:
```{r}
rescale <- function(x) {
  rng <- range(x)
  (x-rng[1]) / (rng[2] - rng[3])
}
```

```{r}
rescale(x)
```

Some more uses of our function:
```{r}
# What will this do?
rescale(c(1,2,NA,3,10))
# How do we fix this?
```

```{r}
x <- c(1,2,NA,3,10)
rng <- range(x)
rng
# How do we fix this?
```

```{r}

```

Let's create a new function and add a new argument to remove our missing values:
```{r}
rescale2 <- function(x) {
  rng <- range(x, na.rm=TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2(c(1,2,NA,3,10))
```

Going too far!
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
  rng <-range(x, na.rm=na.rm)
  print("Hello")
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  print("is it me you are looking for?")
  
  if(plot) {
    plot(answer, typ="b", lwd=4)
  }
  
  print("I can see it in ...")
  return(answer)
}
```

```{r}
rescale3(x, plot=FALSE)
```
```{r}
rescale3(x, plot=TRUE)
```

### Section B
```{r}
library(bio3d)
```
```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Q1. What type of object is returned from the read.pdb() function? 
```{r}
class(s1)
str(s1)
```

Q2. What does the trim.pdb() function do?
```{r}
?trim.pdb()
```

Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?


Q4. What would be a better plot to compare across the different proteins?
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, ty="l", col="blue")
points(s3.b, ty="l", col="red")
```

Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

Q6. How would you generalize the original code above to work with any set of input protein structures?
*For homework*